---
title: "Lecture13"
author: "Andrew Valadez"
date: "May 16, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 1


Q1: What are those 4 candidate SNPs?
[HINT, you will may want to check the first few links of search result]
rs12936231, rs8067378, rs9303277, and rs7216389 

Q2: What three genes do these variants overlap or effect?
[HINT, you can find the information from the ENSEMBLE page as shown in the image
below with red rectangles]
ZPBP2, GSDMB, and ORMDL3


https://uswest.ensembl.org/Homo_sapiens/Variation/Explore?r=17:39872367-39873367;v=rs12936231;vdb=variation;vf=7949754


#1000 Genome Project data

Read 1000 genome data for MXL dataset


Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL)
are homozygous for the asthma associated SNP (G|G)?
[HINT: You can download a CVS file for this population from ENSEMBLE and use the R
functions read.csv(), and table() to answer this question]
14% 


```{r}
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
genotype
```


```{r}
  table(genotype$Genotype..forward.strand.)/nrow(genotype)
```




#Base quality scores from fastqsanger

```{r}
library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@"))
phred
```





#Population scale analysis


```{r}
geno <- read.table("rs8067378_ENSG00000172057.6.txt")
head(geno)
summary(geno)
```

```{r}
summary(geno$exp[geno$geno == "G/G"])
summary(geno$exp[geno$geno == "A/G"])
summary(geno$exp[geno$geno == "A/A"])

```



```{r}
boxplot(exp ~ geno, data=geno, notch=TRUE)
```

```{r}
library(ggplot2)
```

```{r}
ggplot(geno, aes(geno, exp)) +geom_boxplot()
```



