---
title: "Lecture15"
author: "Andrew Valadez"
date: "May 24, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source("https://bioconductor.org/biocLite.R")

biocLite("gage")

```



```{r}
library(DESeq2)
```


```{r}
metaFile <- "data/GSE37704_metadata.csv"
countFile <- "data/GSE37704_featurecounts.csv"


```

 Import metadata and take a peak


```{r}
colData= read.csv(metaFile,row.names = 1)
head(colData)
```
 Import countdata

```{r}
countData = read.csv(countFile, row.names=1)
head(countData)
```


```{r}
# Note we need to remove the odd first $length col
countData <- as.matrix(countData[,-1])
head(countData)
```


This looks better but there are lots of zero entries in there so let’s get rid of them as we have no data for these.

 Filter count data where you have 0 read count across all samples.



```{r}
countData = countData[rowSums(countData)>1, ]
head(countData)
```



Nice now lets setup the DESeqDataSet object required for the DESeq() function and then run the DESeq pipeline. This is again similar to our last days hands-on session.

```{r}
dds = DESeqDataSetFromMatrix(countData=countData,
                             colData=colData,
                             design=~condition)
dds = DESeq(dds)
```
Extracting our results table

```{r}
res <- results(dds)
res
```


```{r}
mcols(res, use.names = TRUE)

```


```{r}
summary(res)
```

In the summary of our results printed above (and by default) the FDR level is set to 10% (i.e. adjusted p-value < 0.1) and the log2 fold change threshold is set to 0. Use the alpha and lfcThreshold input arguments to the results() function to change these to an FDR of 5% and a log2 fold change threshold of 2. Then use the summary() function to find out how many genes are up and down at these thresholds.



```{r}
res <- results(dds, lfcThreshold = 2, alpha = 0.05)
mcols(res, use.names = TRUE)
summary(res)

```


or do it like a table
```{r}
table( res$padj < 0.05, res$log2FoldChange > 2)
```


