---
title: "Class11"
author: "Andrew Valadez"
date: "May 9, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section11

https://bioboot.github.io/bimm143_S18/class-material/lecture11-BIMM143_S18.pdf

Q1: What proportion of PDB entries does X-ray crystallography account for? What proportion of
structures are protein?

 125421/140109
[1] 0.8951673

Q2: Type HIV in the search box on the home page and determine how many HIV-1 protease
structures are in the current PDB?

118 Protease structures


oh look it has a csv file
```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)

```
now lets get do the math with the csv data
```{r}
percent <- (p$Total/ sum(p$Total)) * 100
names(percent) <-  row.names(p)
percent
```

when uploading to github and if you want it to include the rmd in a way where it looks prettier, click the gear icon then go to output options, keep markdown source file. ok. now knit again this will make the .md intermediate folder and then you can go in ahead and commit and push onto git. 


##Using Bio3d
```{r}
library(bio3d)
```

Read in our  HIV-Pr structure

```{r}
pdb <- read.pdb("1hsg")
pdb
```



Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein
residues? 
198


Note that the attributes (+ attr:) of this object are listed on the last couple of lines. To find the
attributes of any such object you can use
```{r}
attributes(pdb)
```

Print a subset of $atom data for the first two atoms

```{r}
head(pdb$atom)
```
 Note that individual $atom records can also be accessed like this
```{r}
pdb$atom[1:2, c("eleno","elety", "x","y","z")]
```
Which allows us to do the following
J
```{r}
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab="B-factor")

```


Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a
useful summery of any R object.

```{r}
str(pdb$atom)
```
It is a data frame

 Print a summary of the coordinate data in $xyz
```{r}
pdb$xyz
```

Examine row and column dimensions
```{r}
dim(pdb$xyz)
```

Print coordinatesfor the first two atoms
```{r}
pdb$xyz[1,atom2xyz(1:2)]
```

The Bio3D atom.select() function is arguably one of the most challenging for newcomers to
master. It is however central to PDB structure manipulation and analysis. At its most basic, this
function operates on PDB structure objects (as created by read.pdb()) and returns the
numeric indices of a selected atom subset. These indices can then be used to access the
$atom and $xyz attributes of PDB structure related objects.
For example to select the indices for all C-alpha atoms we can use the following command:



Select all c alpha atoms returnin their indices
```{r}
ca.inds <- atom.select(pdb,"calpha")
ca.inds 
```

Print details of first few atoms
```{r}
head(pdb$atom[ca.inds$atom, ])
```

and the xyz coordinates
```{r}
head(pdb$xyz[,ca.inds$xyz])
```


In addition to the common selection strings (such as ‘calpha’ ‘cbeta’ ‘backbone’ ‘protein’
‘notprotein’ ‘ligand’ ‘water’ ‘notwater’ ‘h’ and ‘noh’) various individual atom properties can be
used for selection



select chain a
```{r}
a.inds <- atom.select(pdb,chain = "A")
```

select c alphas of chain a

```{r}
ca.inds <- atom.select(pdb,"calpha", chain = "A")
```

combine multiple criteria and return where they intersect

```{r}
cab.inds <- atom.select(pdb, elety = c("CA","CB"), chain = "A", resno = 10:20)
```
Q8. Use the Bio3D write.pdb() function to write out a new C-alpha atom only PDB file for
viewing in VMD.

```{r}
inds.ligand <- atom.select(pdb,"ligand")
inds.protein <- atom.select(pdb,"protein")
inds.protein
```
check we have what we want

```{r}
#pdb$atom[inds.ligand$atom]
```


```{r}
#head(pdb$atom[inds.protein$atom])
```

```{r}
#pdb.protein <- trim.pdb(pbd, inds = inds.protein)
pdb.ligand <- trim.pdb(pdb, inds = inds.ligand)
pdb.ligand
```
```{r}
write.pdb(pdb.ligand,file="1hsvligand.pdb")
```


#Section 5 working with multiple PDB files


Download some example PDB files
```{r}
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)

```

```{r}
pdbs <- pdbaln(files, fit = TRUE)
```

Summary

```{r}
pdbs
```

Here the returned object is of class pdbs. Note that it contains a xyz numeric matrix of aligned
C-alpha coordinates, a ali matrix of aligned residues, and a resno matrix of aligned residue
numbers (see the list of associated attributes (+ attr)). These attirbutes can be accessed
using the common $ syntax in R. E.g. use pdbs$ali to access the alignment. To access the
first few rows of the alignment matrix we use standard subsetting syntax for matrices in R:

access first 5 rows and 8 columns
```{r}
pdbs$ali[1:5, 1:8]
```

Associated residues numbers

```{r}
pdbs$resno[1:5, 1:8]
```

inspect alignment for gaps
```{r}
gaps <- gap.inspect(pdbs$ali)
```
indices of non gap columns
```{r}
head(gaps$f.inds)
```

access non gap columns
```{r}
pdbs$ali[,gaps$f.inds]
```

Having the generated pdbs object at hand facilitates a range of possibilities for protein structure
analysis. This includes sequence identity/similarity, structural deviation, rigid core identification
as well as principal component and normal mode analysis. Several Bio3D function are
specifically designed to operate on the pdbs object, including functions seqidentity(),
rmsd(), pca(), core.find(), nma() and many others.
Below we calculate the pairwise sequence identity between the structures of the pdbs
ensemble followed by the root mean square deviation (RMSD):


calculate sequence identity
```{r}
seqidentity(pdbs)
```

calculate rmsd
```{r}
#rmsd(pdbs)
```

calculate rmsd
```{r}
#rd <- rmsd(pdbs)
```







# Section example aplication of adenylate kinase
```{r}
aa = get.seq("1ake_A")

```

blast or hmmr search
```{r}
b <- blast.pdb(aa)
```

plot a summary of search results
```{r}
hits <- plot(b)
```

using only top scoring hits
```{r}
head(hits$pdb.id)
```


fetch pdbs
```{r}
files5 <- get.pdb(hits$pdb.id, path="pdbs", split = TRUE, gzip = TRUE)
```

align structures
```{r}
pdbs5 <- pdbaln(files5)
```

vector containing pdb codes
```{r}
#ids5 <- basename.pdb(pdbs5$ids)
```

draw an alignmen

```{r}
plot(pdbs5, labels=ids)
```
sequence conservation analysis
```{r}
cons <- conserv(pdbs5, method = "entropy22")
```

sse annotations
```{r}
sse <- pdbs2sse(pdbs, ind=1, rm.gaps = FALSE)
```

plot conservaion per residue


```{r}
plotb3(cons, sse=sse, ylab="Sequence entropy")
```

PCA

```{r}
core <- core.find(pdbs5)
```
superimpose all structures to core

```{r}
pdbs5$xyz= pdbfit(pdbs5,core)
```
perform pca
```{r}
pc.xray <- pca(pdbs5)
plot(pc.xray)
```

visualize first principal component
```{r}
mktrj(pc.xray,pc=1,file="pc_1.pdb")
```

